---
title: "Assignment 4 - Problem 2"
author: "Sharif Haason"
date: "`r Sys.Date()`"
output: html_document
---
# Initialize Libraries/Data

```{r setup}
library("Sim.DiffProc")
library(knitr)

#setwd("C:/Users/sharif/Code/fe2/fe621")
setwd("/home/sharif/Code/fe2/fe621")
data <- read.csv("sample_data25.csv")
```

# Set Expressions

These are the models that we are trying to fit.

```{r setexpr}
f1 <- expression(theta[1] * x)
g1 <- expression(theta[2] * x^theta[3])

f2 <- expression(theta[1] + theta[2] * x)
g2 <- expression(theta[3] * x^theta[4])

f3 <- expression(theta[1] + theta[2] * x)
g3 <- expression(theta[3] * sqrt(x))

f4 <- expression(theta[1])
g4 <- expression(theta[2] * x^theta[3])

f5 <- expression(theta[1] * x)
g5 <- expression((theta[2] + theta[3] * x^theta[4]))
```

# Stock 1

For stock 1, model 4 is determined to have the best fit.

```{r stock1}
fit1 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f1, diffusion = g1, start = list(theta1=0.03, theta2=0.19, theta3=0.5), pmle="euler")
coef(fit1)
logLik(fit1)
AIC(fit1)
BIC(fit1)

fit2 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f2, diffusion = g2, start = list(theta1=0.01, theta2=0.02, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit2)
logLik(fit2)
AIC(fit2)
BIC(fit2)

fit3 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=1, theta3=0.5), pmle="euler")
coef(fit3)
logLik(fit3)
AIC(fit3)
BIC(fit3)

fit4 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="euler")
coef(fit4)
logLik(fit4)
AIC(fit4)
BIC(fit4)

fit5 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f5, diffusion = g5, start = list(theta1=0.03, theta2=0.01, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit5)
logLik(fit5)
AIC(fit5)
BIC(fit5)

aic_values <- c(
    fit1 = AIC(fit1),
    fit2 = AIC(fit2),
    fit3 = AIC(fit3),
    fit4 = AIC(fit4),
    fit5 = AIC(fit5)
)

aic_values
best_fit <- names(which.min(aic_values))
cat("The model with the lowest AIC is:", best_fit, "with AIC =", min(aic_values), "\n")
```

# Stock 2

For stock 2, model 3 is determined to have the best fit.

```{r stock2}
fit1 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f1, diffusion = g1, start = list(theta1=0.007, theta2=0.17, theta3=0.5), pmle="euler")
coef(fit1)
logLik(fit1)
AIC(fit1)
BIC(fit1)

fit2 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f2, diffusion = g2, start = list(theta1=0.01, theta2=0.006, theta3=0.15, theta4=0.5), pmle="euler")
coef(fit2)
logLik(fit2)
AIC(fit2)
BIC(fit2)

fit3 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.001, theta3=0.16), pmle="euler")
coef(fit3)
logLik(fit3)
AIC(fit3)
BIC(fit3)

fit4 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.6, theta2=0.16, theta3=0.5), pmle="euler")
coef(fit4)
logLik(fit4)
AIC(fit4)
BIC(fit4)

fit5 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f5, diffusion = g5, start = list(theta1=0.006, theta2=0.007, theta3=0.15, theta4=0.5), pmle="euler")
coef(fit5)
logLik(fit5)
AIC(fit5)
BIC(fit5)

aic_values <- c(
    fit1 = AIC(fit1),
    fit2 = AIC(fit2),
    fit3 = AIC(fit3),
    fit4 = AIC(fit4),
    fit5 = AIC(fit5)
)

aic_values
best_fit <- names(which.min(aic_values))
cat("The model with the lowest AIC is:", best_fit, "with AIC =", min(aic_values), "\n")
```

# Stock 3

For stock 3, model 4 is determined to have the best fit.

```{r stock3}
fit1 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f1, diffusion = g1, start = list(theta1=0.007, theta2=0.07, theta3=0.4), pmle="euler")
coef(fit1)
logLik(fit1)
AIC(fit1)
BIC(fit1)

fit2 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f2, diffusion = g2, start = list(theta1=0.01, theta2=0.007, theta3=0.06, theta4=0.43), pmle="euler")
coef(fit2)
logLik(fit2)
AIC(fit2)
BIC(fit2)

fit3 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=-0.005, theta2=0.007, theta3=0.045), pmle="euler")
coef(fit3)
logLik(fit3)
AIC(fit3)
BIC(fit3)

fit4 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.75, theta2=0.77, theta3=-0.11), pmle="euler")
coef(fit4)
logLik(fit4)
AIC(fit4)
BIC(fit4)

fit5 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f5, diffusion = g5, start = list(theta1=0.007, theta2=0.01, theta3=0.06, theta4=0.43), pmle="euler")
coef(fit5)
logLik(fit5)
AIC(fit5)
BIC(fit5)

aic_values <- c(
    fit1 = AIC(fit1),
    fit2 = AIC(fit2),
    fit3 = AIC(fit3),
    fit4 = AIC(fit4),
    fit5 = AIC(fit5)
)

aic_values
best_fit <- names(which.min(aic_values))
cat("The model with the lowest AIC is:", best_fit, "with AIC =", min(aic_values), "\n")
```

# Stock 4

For stock 4, model 4 is determined to have the best fit.

```{r stock4}
fit1 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f1, diffusion = g1, start = list(theta1=0.03, theta2=0.19, theta3=0.5), pmle="euler")
coef(fit1)
logLik(fit1)
AIC(fit1)
BIC(fit1)

fit2 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f2, diffusion = g2, start = list(theta1=0.01, theta2=0.02, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit2)
logLik(fit2)
AIC(fit2)
BIC(fit2)

fit3 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=0.01, theta3=1), pmle="euler")
coef(fit3)
logLik(fit3)
AIC(fit3)
BIC(fit3)

fit4 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="euler")
coef(fit4)
logLik(fit4)
AIC(fit4)
BIC(fit4)

fit5 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f5, diffusion = g5, start = list(theta1=0.03, theta2=0.01, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit5)
logLik(fit5)
AIC(fit5)
BIC(fit5)

aic_values <- c(
    fit1 = AIC(fit1),
    fit2 = AIC(fit2),
    fit3 = AIC(fit3),
    fit4 = AIC(fit4),
    fit5 = AIC(fit5)
)

aic_values
best_fit <- names(which.min(aic_values))
cat("The model with the lowest AIC is:", best_fit, "with AIC =", min(aic_values), "\n")
```

# Stock 5

For stock 5, model 3 is determined to have the best fit.

```{r stock5}
fit1 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f1, diffusion = g1, start = list(theta1=0.03, theta2=0.19, theta3=0.5), pmle="euler")
coef(fit1)
logLik(fit1)
AIC(fit1)
BIC(fit1)

fit2 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f2, diffusion = g2, start = list(theta1=0.01, theta2=0.02, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit2)
logLik(fit2)
AIC(fit2)
BIC(fit2)

fit3 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=0.1, theta3=0.5), pmle="euler")
coef(fit3)
logLik(fit3)
AIC(fit3)
BIC(fit3)

fit4 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="euler")
coef(fit4)
logLik(fit4)
AIC(fit4)
BIC(fit4)

fit5 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f5, diffusion = g5, start = list(theta1=0.03, theta2=0.01, theta3=0.19, theta4=0.54), pmle="euler")
coef(fit5)
logLik(fit5)
AIC(fit5)
BIC(fit5)

aic_values <- c(
    fit1 = AIC(fit1),
    fit2 = AIC(fit2),
    fit3 = AIC(fit3),
    fit4 = AIC(fit4),
    fit5 = AIC(fit5)
)

aic_values
best_fit <- names(which.min(aic_values))
cat("The model with the lowest AIC is:", best_fit, "with AIC =", min(aic_values), "\n")
```

# Parametrization

```{r euler}
fit1 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="euler")
fit2 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.01, theta3=0.16), pmle="euler")
fit3 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.75, theta2=0.77, theta3=-0.11), pmle="euler")
fit4 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="euler")
fit5 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.0001, theta3=1.5), pmle="euler")
fits <- list(fit1, fit2, fit3, fit4, fit5)
coefs <- sapply(fits, coef)
results <- data.frame(
    Stock = c("Stock1", "Stock2", "Stock3", "Stock4", "Stock5"),
    Model = c("4", "3", "4", "4", "3"),
    Theta1 = coefs["theta1", ],
    Theta2 = coefs["theta2", ],
    Theta3 = coefs["theta3", ]
)

# Display the table
kable(results, caption = "Parameter Estimates and Models for Each Stock (Euler)")
```

```{r ozaki}
fit1 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="ozaki")
fit2 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.001, theta3=0.16), pmle="ozaki")
fit3 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.75, theta2=0.77, theta3=-0.11), pmle="ozaki")
fit4 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="ozaki")
fit5 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=0.1, theta3=0.5), pmle="ozaki")
fits <- list(fit1, fit2, fit3, fit4, fit5)
coefs <- sapply(fits, coef)
results <- data.frame(
    Stock = c("Stock1", "Stock2", "Stock3", "Stock4", "Stock5"),
    Model = c("4", "3", "4", "4", "3"),
    Theta1 = coefs["theta1", ],
    Theta2 = coefs["theta2", ],
    Theta3 = coefs["theta3", ]
)

# Display the table
kable(results, caption = "Parameter Estimates and Models for Each Stock (Ozaki)")
```

```{r shoji-ozaki}
fit1 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="shoji")
fit2 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.001, theta3=0.16), pmle="shoji")
fit3 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.75, theta2=0.77, theta3=-0.11), pmle="shoji")
fit4 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="shoji")
fit5 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=0.1, theta3=0.5), pmle="shoji")
fits <- list(fit1, fit2, fit3, fit4, fit5)
coefs <- sapply(fits, coef)
results <- data.frame(
    Stock = c("Stock1", "Stock2", "Stock3", "Stock4", "Stock5"),
    Model = c("4", "3", "4", "4", "3"),
    Theta1 = coefs["theta1", ],
    Theta2 = coefs["theta2", ],
    Theta3 = coefs["theta3", ]
)

# Display the table
kable(results, caption = "Parameter Estimates and Models for Each Stock (Shoji-Ozaki)")
```

```{r kessler}
fit1 <- fitsde(data = ts(data$stock1, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="kessler")
fit2 <- fitsde(data = ts(data$stock2, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=0.5, theta2=0.001, theta3=0.16), pmle="kessler")
fit3 <- fitsde(data = ts(data$stock3, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.75, theta2=0.77, theta3=-0.11), pmle="kessler")
fit4 <- fitsde(data = ts(data$stock4, frequency = 365), drift = f4, diffusion = g4, start = list(theta1=0.03, theta2=0.19, theta3=0.54), pmle="kessler")
fit5 <- fitsde(data = ts(data$stock5, frequency = 365), drift = f3, diffusion = g3, start = list(theta1=1, theta2=0.1, theta3=0.5), pmle="kessler")
fits <- list(fit1, fit2, fit3, fit4, fit5)
coefs <- sapply(fits, coef)
results <- data.frame(
    Stock = c("Stock1", "Stock2", "Stock3", "Stock4", "Stock5"),
    Model = c("4", "3", "4", "4", "3"),
    Theta1 = coefs["theta1", ],
    Theta2 = coefs["theta2", ],
    Theta3 = coefs["theta3", ]
)

# Display the table
kable(results, caption = "Parameter Estimates and Models for Each Stock (Kessler)")
```

I decided that the Kessler model had the best fit, since we're calibrating to real data, it's able to directly model better.
